---
title: "Lab 5 Fun Times"
author: "Annette Hilton"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives: 

- Getting counts for different groups
- Use {lubridate} to parse dates
- Find confidence intervals and do t-tests with t.test() 
- Heatmap with geom_tile()


```{r, include = FALSE}
 # Attach packages

library(tidyverse)
library(here)
library(janitor)

```

```{r}
# Import lobster data 
# Put in clarification for NA values 

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"), 
                              na = "-99999") %>% 
  clean_names()
  
```

Use the tidyr::uncount() function to expand a frequency variable (lobster count)

```{r}

lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)
```

Now each lobster has its own row, this is tidy format 

Let's look at the data in various visualizations 

```{r}
# Exploratory data visualization 

# Jitter plot 
ggplot(data = lobster_tidy, 
       aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site), 
              width = 0.2)

# Violin plot 
ggplot(data = lobster_tidy, 
       aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site))

# Histogram 
ggplot(data = lobster_tidy, 
       aes(x = size_mm)) +
  geom_histogram (aes( fill = site)) +
  facet_wrap(~site, scale = "free") #function to separate histogram into each site, have scale adjust for each site (max and min for each group) 

# Quantile-quantile plot 
ggplot(data = lobster_tidy, 
       aes(sample = size_mm)) +
  geom_qq(size = 0.5) +
  facet_wrap(~site, scale = "free")

```

Use the lubridate package to parse dates and times

Add a new column with mutate(), that contains my date as an actual date 

```{r}
# Change from a character to a date class 
# Create new column with dates in ISO format (most accepted format for dates)

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date))

```

Now we will parse 'date' column to get different pieces (month, year) separated 

```{r}
# p.s- lubridate only works on a column that is classed as 'date' 

lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE), 
    obs_year = lubridate::year(date_new)) # separate month from the date column, label allows you to have it be month abbrev. (Aug) rather than number 

```

Let's ask some questions about counts of the lobsters. Let's count by different groupings

Let's count the number of lobsters by year and month: 

```{r}
# The count function is AMAZING, it's essentially an easy replacement to 'group_by' 

lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

lobster_ysite <- lobster_parse_date %>% 
  count(obs_year, site)

lobster_site <- lobster_parse_date %>% 
  count(site)
```

```{r}
# BUT what if instead for each site, I want to calculate statistics for each site and have it in the same table 
# revert to using group_by and summarize and n()

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE), # calculate mean 
    sd_size = sd(size_mm, na.rm = TRUE), # calculate standard deviation 
    sample_n = n() # find sample size 
    )
```

Confidence intervals and t-tests 

Use the t.test() function to find confidence intervals and perform t-tests

```{r}
# Let's find the confidence interval for lobster sizes at Isla Vista Beach 
# I will create a vector with those observations 

ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% # filter by IVEE 
  pull(size_mm) # pulls out size column only, creates vector of observations 

# perform a one-sample t-test 

t.test(ivee_lobsters)
```

Two-sample t-test 

Is there a significant difference in mean lobster lengths between Naples Reef and Mohawk Reef? 

```{r}
# Two sample ttest 
# Make vectors of each sample first (naples, mohawk) 

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(napl_sample, mohk_sample)
mohk_napl_ttest
```

Make a statement about this statistical hypothesis test: 

Mean lobster size differed significantly between Mohawk and Naples Reefs (t(`r mohk_napl_ttest$parameter`) = 19.849)

Maybe an easier way to do a two-sample ttest

```{r}
# create data frame with two groups I want to compare 

lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

# Run t-test 

ttest2 <- t.test(size_mm ~ site, data = lobster_2sample)
ttest2
```

Make a geom_tile heatmap 

```{r}

ggplot(data = lobster_ysite, 
       aes(x = obs_year, 
           y = site)) + 
  geom_tile(aes(fill = n))
```

